---
title: "Missouri Headwaters Basin Conservation Prioritization Tool"
#author: "Meredith McClure"
date: "last updated 12/14/2017"
output: html_notebook
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
Use this interactive document to set up the Zonation spatial prioritization tool run. Here you will select your conservation action of interest, assign weights to conservation targets that are linked to that action, and set additional run parameters.  

When you are finished, press the 'Save settings' button. This will save the your settings to files needed by the Zonation tool. 

***

### Prioritization Settings
```{r include=FALSE}
library(raster, quietly=T)
library(rgdal, quietly=T)
library(leaflet, quietly=T)
library(zonator, quietly=T)
```

```{r globalfiles, echo=FALSE}
targ.summ<-read.csv("data/actiontargets.csv")
act.lyrs<-read.csv("data/actionlayers.csv")
mhb <- shapefile("MissouriHeadwaters_HUC6/MissouriHeadwaters_HUC6.shp")
huc12 <- shapefile("MissouriHeadwaters_HUC12/MissouriHeadwaters_HUC12.shp")
```

```{r eruptions, echo=FALSE, warning=TRUE}
fluidPage(sidebarLayout(
  sidebarPanel(
    wellPanel(
      textInput("run.name","Give your Zonation run a name:","e.g., 'MyAction_Run1'"),
  
      radioButtons("action.sel", "Select Action of Interest:",
               c("Conservation easement" = "option1",
                 "Retrofit road culverts/bridges" = "option2",
                 "Road decommissioning" = "option3",
                 "Grazing management" = "option4",
                 "Adjust irrigation practices" = "option5",
                 "Soil health management" = "option6",
                 "Stream/riparian restoration" = "option7",
                 "Thinning/burning" = "option8"
                 ))
    ),
  
    wellPanel(
      h5("Save your settings for Zonation, or load your last-used settings:"),
          actionButton("save.inputs", "Save settings"),
          actionButton("load.inputs", "Load settings") 
    ),
    
    wellPanel(
      textOutput("instruc")
    )
  ),
  
  mainPanel(
      wellPanel(
          h5("Set weights for each target:"),
          uiOutput("wt.sel")
      ),
  
      wellPanel(h5("Choose removal rule:"),
        radioButtons("rem.rule","",c("Basic Core-Area Zonation"="1","Additive Benefit Function"="2","Target-Based Planning"="3","Generalized Benefit Function"="4","Random Removal"="5"),selected="1")
      )
  )
))

output$wt.sel <- renderUI({
      targets<-targ.summ[targ.summ[,input$action.sel]==1,"name"]
      lapply(1:length(targets), function(i) {
        if(length(input$action.sel)==0) {output$wt.sel0 <- renderText({" "})} else {
        sliderInput(inputId=paste0("targ",i),label=targets[i],min=0,max=1,value=0,step=0.1)}
        })
})

observeEvent(input$load.inputs,{   
    
    if(!file.exists('runs/lastinputs.RDS')) {return(NULL)}
    
    savedInputs <- readRDS('runs/lastinputs.RDS')
    
    inputIDs      <- names(savedInputs) 
    inputvalues   <- unlist(savedInputs) 
    for (i in 1:length(savedInputs)) { 
      session$sendInputMessage(inputIDs[i],  list(value=inputvalues[[i]]) )
    }
  })
  
  observeEvent(input$save.inputs,{ 
    saveRDS(reactiveValuesToList(input), file = "runs/lastinputs.RDS")

    project_path<-file.path("runs")
    variant_name<-input$run.name
    input_raster_dir<-"data/01_biodiversity_features/02_habitats/forest"
    create_zproject(name=input$run.name,dir=project_path,variants=variant_name,
                spp_template_dir=input_raster_dir,spp_file_pattern="*.tif",overwrite=T,
                weight=c(input$targ1,input$targ2,input$targ3,input$targ4,input$targ5))

    dat<-list("Settings" = list("removal rule"=input$rem.rule, "warp factor"=1, "edge removal"=1, 
                            "use planning unit layer"=1, "planning unit layer file" = "",
                            "use mask"=1, "mask file"=act.lyrs[act.lyrs[,'label']==input$action.sel,"masklyr"],
                            "use tree connectivity"=1,"tree connectivity file"=""))
    write_dat(dat,paste0(project_path,"/",input$run.name,"/",variant_name,"/",variant_name,".dat"),overwrite=T)
    output$instruc<-renderText({paste("Your run settings have been saved to a '.bat' file in ",project_path,"/",input$run.name,". Next, open the Zonation tool, load your file, and hit 'Run'.",sep="")})
  })  
```

***

This tool was developed by [The Center for Large Landscape Conservation](http://www.largelandscapes.org) and [Conservation Science Partners](http://www.csp-inc.org).  
This project was funded by [The National Fish & Wildlife Foundation](http://www.nfwf.org).



