---
title: "Missouri Headwaters Basin Cross-Realm Management Prioritization Tool"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
runtime: shiny
---

```{r}
#####################################################################
#                                                                   #
#                                                                   #
#  TO RUN THIS TOOL, CLICK 'Run Document' NEXT TO THE GREEN ARROW   #
#             DIRECTLY ABOVE IN THE HEADER OF THIS TAB.             #
#                                                                   #
#  If you are running this tool for the first time, make sure that  #
#   you have run 'MHB_ToolSetup.R' first to install the necessary   #
#                          R packages.                              #
#                                                                   #
#          (There is no need to interact with the code              #
#               below - please do not modify it!)                   #
#                                                                   #
#                                                                   #
#####################################################################
```

<style>                     
.navbar {
  background-color:#2F4F4F;
  border-color:white;
}
.navbar-brand {
color:white!important;
}
</style>  

```{r global, include=FALSE}
#Load packages
packages<-c("rmarkdown","raster","rgdal","rgeos","leaflet","mapview","RColorBrewer","zonator")
lapply(packages,library,character.only=T)  
source('create_zproject_rev.R')  

#Load common data
targ.summ<-read.csv("runs/data/actiontargets.csv",header=T)
act.lyrs<-read.csv("runs/data/actionlayers.csv",header=T)
mhb <- shapefile("runs/data/MissouriHeadwaters_HUC6/MissouriHeadwaters_HUC6.shp")
huc12 <- shapefile("runs/data/MissouriHeadwaters_HUC12/MissouriHeadwaters_HUC12_attr.shp")
```

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r userInputs}
    #USER INPUT (TEXT): NAME RUN
    textInput("run.name","Name your prioritization run:","MyAction_Run1")
  
    #USER INPUT (DROPDOWN MENU): ACTION SELECTION
    selectInput("action.sel", "Select Action of Interest:",
                    c("Private land protection" = "option1",
                     "Upgrade culverts/bridges" = "option2",
                     "Decommission Roads" = "option3",
                     "Reduce grazing pressure" = "option4",
                     "Adjust irrigation practices" = "option5",
                     "Manage soil health" = "option6",
                     "Stream/riparian restoration" = "option7",
                     "Fuels management" = "option8",
                     "Encroachment control" = "option9")
                  )
    
    #REACTIVE USER INPUT (SLIDERS): WEIGHT SELECTION
    h5("Set weights for each target:")
    uiOutput("wt.opt")
        
      #REACTIVELY RENDER SLIDERS: WEIGHT OPTIONS DEPENDENT ON ACTION SELECTION
      output$wt.opt <- renderUI({
        targets<<-targ.summ[targ.summ[,input$action.sel]==1,"name"]
        if(length(input$action.sel)>0) {
          lapply(1:length(targets), function(i) {
              sliderInput(inputId=paste0("targ",i),label=targets[i],min=0,max=1,value=0,step=0.1)}
            )}
        })
      
    #REACTIVE USER INPUT (DROPDOWN MENU): DIRECTIONAL CONNECTIVITY SETTING
    uiOutput("dirconn.opt")

      #REACTIVELY RENDER MENU: DIRECTIONAL CONNECTIVITY OPTIONS DEPENDENT ON APPLICABLE TARGETS
      observeEvent(input$action.sel,{
          output$dirconn.opt <-renderUI({
            if(sum(targets %in% c("Stream health","Normative flow regime"))>0) {
              selectInput("dirconn.sel","Consider downstream benefits for stream target(s)?",
                    c("No benefits" = "1",
                     "Small benefits" = "2",
                     "Moderate benefits" = "3",
                     "Large benefits" = "4")
                    )}
          })
      })
```

Row {.tabset data-height=850}
-----------------------------------------------------------------------

### Explore the Map
    
```{r mapInterface}
    #CREATE MAP OBJECT
    tags$style(type = "text/css", "#map {height: calc(100vh) !important;}")
    leafletOutput('map', width=800, height=600)
    #bounds<-extent(-113.9473,-109.5527,44.24289,46.55711)
    bounds<-extent(-112.8486,-110.6514,44.82144,45.97856)

    #RENDER MAP: BASE
    output$map <- renderLeaflet({
      leaflet(mhb,options = leafletOptions(minZoom = 7)) %>%
        addProviderTiles("Esri.WorldTopoMap", group="Topographical") %>%
        addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
        addPolygons(data=huc12, fill = FALSE, color="#404040", weight = 1, opacity = 1, group="Boundaries") %>%
        addPolygons(fill = FALSE, color = "black", weight = 3, opacity = 1, group="Boundaries") %>%
        setView(lng = -111.75, lat = 45.4,  zoom = 8) %>%
        addHomeButton(ext=bounds, layer.name = "Full Extent", position = "topleft", add = TRUE)
    })
       
    #REACTIVELY ADD MAP LAYERS: CONDITION LAYERS DEPENDENT ON ACTION SELECTION
    pal <- colorNumeric(palette = c("white","darkslategray"), domain = 0:1)

    observeEvent(input$action.sel, ({
        targetIDs<-targ.summ[targ.summ[,input$action.sel]==1,"abbv"]
        targetNames<<-as.vector(targ.summ[targ.summ[,input$action.sel]==1,"name"]) 
        maskLyr<-shapefile(paste0("runs/data/ActionMasks/",act.lyrs[act.lyrs$label==input$action.sel,"masklyr_shp"]))

        targCond<-list()
        for (i in 1:length(targetIDs)) {
          targCond[[i]]<-huc12[,as.character(targetIDs[i])]
          names(targCond[[i]])<-"targ"
        }
    
        leafletProxy("map") %>%
            clearControls() %>%
            clearGroup(c("Action Footprint",targ.summ[targ.summ[,"option1"]==1,"name"])) %>%
            hideGroup(c("Action Footprint",targetNames)) %>%
            addTiles() %>%
            {
              for(i in 1:length(targCond)){
                . <- addPolygons(.,data=targCond[[i]], color = "#404040", weight = 1, opacity = 1, 
                                 fillOpacity = 0.75, fillColor = ~pal(targ), group=targetNames[i])
              }
              return(.)
            } %>%
            addPolygons(data = maskLyr, stroke = FALSE, fill = TRUE,  
                        fillColor = "black", fillOpacity = 0.75, group = "Action Footprint") %>%            
            addLayersControl(baseGroups = c("Topographical", "Satellite"),
                            overlayGroups = c("Action Footprint",targetNames),       
                            options = layersControlOptions(collapsed = FALSE)) %>%
            showGroup(c(targetNames[1])) %>%
            addLegend("bottomright", pal = colorNumeric(pal,domain=0:1,reverse=T), values=0:1, opacity = 0.75, 
                      labFormat = function(type, cuts, p) { 
                          n = length(cuts) 
                          cuts[1] = "High" 
                          for (i in 2:(n-1)) {cuts[i] = ""} 
                          cuts[n] = "Low" 
                          paste0(cuts[-n], cuts[-1])}, 
                      title = "Conservation Target <br> Opportunity Value")
      })
      )

    #REACTIVELY ADD MAP LAYERS: ZONATION OUTPUT DEPENDENT ON CHECK STATUS
    observeEvent(input$check.stat,{
      project_path<-file.path("runs")
      variant_name<-input$run.name
    
      outputFile<-list.files(path = paste0("./runs/",variant_name), pattern = "rank.compressed.tif",full.names=T,recursive=T) 
      if (isTRUE(file.exists(outputFile))) {
        output$stat.done<-renderText({"Your prioritization is done! Please wait a moment for the map to load."})
        
        #Read result and extract result to huc12 polygons by values at centroids
        result<-raster(outputFile)
          huc12c <- suppressWarnings(extract(result, huc12c, sp=TRUE))
          huc12@data <- cbind(huc12@data, huc12c@data[ncol(huc12c@data)])    
          names(huc12@data)[ncol(huc12@data)] <- "result"
          
        rpal<-colorNumeric(palette = "BuPu", domain = 0:1)
          
        leafletProxy("map") %>%
          clearControls() %>%
          addTiles() %>%
          addPolygons(data=huc12, color = "#404040", weight = 1, opacity = 1, 
                                 fillOpacity = 0.75, fillColor = ~rpal(result), group="Prioritization Result") %>%
          addLayersControl(baseGroups = c("Topographical", "Satellite"),
                            overlayGroups = c("Prioritization Result",targetNames,"Action Footprint"),       
                            options = layersControlOptions(collapsed = FALSE)) %>%
          addLegend("bottomright",pal = colorNumeric(rpal,domain=0:1,reverse=T), values = 0:1, opacity = 0.75, 
                    labFormat = function(type, cuts, p) { 
                          n = length(cuts) 
                          cuts[1] = "High" 
                          for (i in 2:(n-1)) {cuts[i] = ""} 
                          cuts[n] = "Low" 
                          paste0(cuts[-n], cuts[-1])},
                    title = "Conservation <br> Priority Rank")
    }
    else{
      output$stat.running<-renderText({"Your prioritization is still running. Please wait."})
    }
    
  })
```

### Get More Information

#### Getting Started

This tool is designed to help identify priority subwatersheds in the Missouri Headwaters Basin where there is potential for a conservation action of interest to provide the greatest net benefits to multiple conservation targets in both terrestrial and freshwater realms. 

Start by selecting your conservation action of interest, then assign weights to conservation targets that are linked to that action, based on their relative importance to you. When the prioritization process is finished, you can use the interactive map to explore the results.

***

#### Resources

For further instructions, more information on the input data and prioritization algorithm, and other supporting information, download the [full documentation](https://drive.google.com/open?id=1gupPNpsWDjBk7jhfqGiyarJwMEDzrjMJ) and/or the [quick-start guide](https://drive.google.com/open?id=1E2ejO9zuzVn1l7yhlpasIN7z9js-hhyC). 

Please see the README file for disclaimer information. 

***

#### Credits

This custom decision support tool was developed by [The Center for Large Landscape Conservation](http://www.largelandscapes.org) and [Conservation Science Partners](http://www.csp-inc.org), with funding from [The National Fish & Wildlife Foundation](http://www.nfwf.org). The prioritization is run using [Zonation 4.0](https://www.helsinki.fi/en/researchgroups/metapopulation-research-centre/software#section-14300) produced by Atte Moilanen and colleagues, and an adaptation of the [zonator](https://github.com/cbig/zonator) package for R developed by Joona Lehtomaki.

Please cite use of this tool, associated data layers, and documentation as: Creech, T., McClure, M., and van Rees, C. 2018. A Conservation Prioritization Tool for the Missouri Headwaters Basin. Center for Large Landscape Conservation, Bozeman, MT.

Contact: [Tyler Creech](tyler@largelandscapes.org), [Meredith McClure](meredith@csp-inc.org)

***


Row {data-height=150}
-----------------------------------------------------------------------

```{r saveRun}
    #USER INPUT: ACTION BUTTONS
    column(3,
        actionButton("load.inputs", "Load last settings", 
            style="background-color: #528B8B; border-color: #2F4F4F")
    )
    column(3,    
        actionButton("save.settings", "Save run settings",
            style="background-color: #528B8B; border-color: #2F4F4F"),
        textOutput("save.confirm")
    )
    column(3,
        actionButton("run.tool", "Run prioritization",
            style="background-color: #528B8B; border-color: #2F4F4F"),
        textOutput("please.wait"),
        textOutput("save.error")
    )
    column(3,
        actionButton("check.stat","Check run status",
            style="background-color: #528B8B; border-color: #2F4F4F"),
        textOutput("stat.running"),
        textOutput("stat.done")
    )
    
    #REACTIVELY LOAD PREVIOUS SETTINGS (LOAD BUTTON)
    observeEvent(input$load.inputs,{   
      if(!file.exists('runs/lastinputs.RDS')) {return(NULL)}
    
      savedInputs <- readRDS('runs/lastinputs.RDS')
      inputIDs      <- names(savedInputs) 
      inputvalues   <- unlist(savedInputs) 
    
      for (i in 1:length(savedInputs)) { 
        session$sendInputMessage(inputIDs[i],  list(value=inputvalues[[i]]))
      }
    })
  
    #REACTIVELY SET UP ZONATION FILES (SAVE BUTTON) 
    observeEvent(input$save.settings,{
      inputList<-reactiveValuesToList(input)
      saveRDS(inputList, file = 'runs/lastinputs.RDS')

      project_path<-file.path("MHB_PrioritizationTool")
      variant_name<-input$run.name
      input_raster_dir<-paste0("runs/data/TargetCondition/",
                               unique(targ.summ[targ.summ[,input$action.sel]==1,"dir"]))   
    
      targWts<-unlist(inputList[grep("targ", names(inputList))])[1:length(targets)]
      targNames<-as.vector(targ.summ[targ.summ[,input$action.sel]==1,"name"])
      targCurves<-ifelse(targNames %in% c("Stream health","Normative flow regime"),as.numeric(input$dirconn.sel),1)
      
      #This uses a revised version of zonator's 'create_zproject' function that correctly aligns subdirectories
      create_zproject_rev(name="runs",dir=".",variants=variant_name,
                        spp_template_dir=input_raster_dir,spp_file_pattern="*.tif",   
                        overwrite=T,
                        weight=targWts, 
                        bqp=targCurves, #upstream curve function
                        bqp_p=1) #downstream curve function 
      
      #See Zonation 4.0 user manual for table of all run settings options
      dat<-list("Settings" = list("removal rule"=2, "edge removal"=1, 
                    "use planning unit layer"=1, 
                    "planning unit layer file"=paste0("runs/data/ZonationData/HUC12_zones_270m_LEAFLET.tif"),  
                    "use cost"=0, 
                    "use mask"=0, 
                    "use tree connectivity"=1, 
                    "tree connectivity file"=paste0("runs/data/ZonationData/MHB_dirconn.txt"),
                    "BQP profiles file" = paste0("runs/data/ZonationData/MHB_NQP.txt")))
      
      write_dat(dat,paste0("./runs/",variant_name,"/",variant_name,".dat"),overwrite=T)
      output$save.confirm <- renderText ({"Your settings have been saved."})
  })  
  
  #REACTIVELY RUN ZONATION FROM SHELL (RUN BUTTON) 
  observeEvent(input$run.tool,{
    project_path<-file.path("runs")
    variant_name<-input$run.name
    
    if (isTRUE(file.exists(paste0("./runs/",variant_name,".bat")))) {
      shell.exec(paste0("runs\\",  
                        variant_name,".bat"))
      huc12c <<- SpatialPointsDataFrame(gCentroid(huc12, byid=TRUE), huc12@data["HUC12"], match.ID=FALSE)
      output$please.wait<-renderText({paste("Please wait. Your results will be saved to 
                                            'runs/",input$run.name,"'.",sep="")})
    }
    else{
      output$save.error <- renderText({"You must first save your run settings"})
    }
  })
  
```