---
title: "Conservation Prioritization Tool: Example with Tutorial Data"
#author: "Meredith McClure"
date: "last updated 01/04/2017"
output: html_notebook
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
**This tool is designed to help identify priority subwatersheds where there is potential for a conservation action of interest to provide the greatest net benefits to multiple conservation targets in the Missouri Headwaters Basin. 
Start by selecting your conservation action of interest, then assign weights to conservation targets that are linked to that action, based on their relative importance to you. When the prioritization process is finished, explore the results using the interactive map and plots.**

***
*This custom decision support tool was developed by [The Center for Large Landscape Conservation](http://www.largelandscapes.org) and [Conservation Science Partners](http://www.csp-inc.org), and was funded by [The National Fish & Wildlife Foundation](http://www.nfwf.org). The prioritization is run using [Zonation 4.0](https://www.helsinki.fi/en/researchgroups/metapopulation-research-centre/software#section-14300) and the [zonator](https://github.com/cbig/zonator) package for R.*

***

### Prioritization Settings

```{r include=FALSE}
if (!require("pacman")) install.packages("pacman",repos = "http://cran.us.r-project.org")
  pacman::p_load(zonator, raster, rgdal, leaflet)
source('E:/Projects/NFWF_IntegratedLI/code/R/NFWF_MHB/create_zproject_rev.R')  
```

```{r globalfiles, echo=FALSE, warning=FALSE}
targ.summ<-read.csv("runs/data/actiontargets.csv",header=T)
act.lyrs<-read.csv("runs/data/actionlayers.csv",header=T)
```

```{r eruptions, echo=FALSE, warning=TRUE}
fluidPage(sidebarLayout(
  sidebarPanel(
    wellPanel(
      textInput("run.name","Give your prioritization a name:","e.g., 'MyAction_Run1'"),
  
      radioButtons("action.sel", "Select Action of Interest:",
               c("Action 1" = "option1",
                 "Action 2" = "option2",
                 "Action 3" = "option3",
                 "Action 4" = "option4")
                 )
    ),
  
    wellPanel(
      actionButton("load.inputs", "Load last-used settings") 
    ),
    
    wellPanel(
      actionButton("save.settings", "Save run settings"),
      textOutput("save.confirm")
    ),
  
    wellPanel(
      actionButton("run.tool", "Run prioritization"),
      textOutput("please.wait"),
      textOutput("save.error")
    )
    #Add 'check status' button: if files don't exist, message, else, show maps/plots
  ),
  
  mainPanel(
      wellPanel(
          h5("Set weights for each target:"),
          uiOutput("wt.sel")
      )
      )
  )
)

output$wt.sel <- renderUI({
      targets<-targ.summ[targ.summ[,input$action.sel]==1,"name"]
      lapply(1:length(targets), function(i) {
        if(length(input$action.sel)==0) {output$wt.sel0 <- renderText({" "})} else {
        sliderInput(inputId=paste0("targ",i),label=targets[i],min=0,max=1,value=0,step=0.1)}
        })
})

observeEvent(input$load.inputs,{   
    if(!file.exists('runs/lastinputs.RDS')) {return(NULL)}
    
    savedInputs <- readRDS('runs/lastinputs.RDS')
    inputIDs      <- names(savedInputs) 
    inputvalues   <- unlist(savedInputs) 
    
    for (i in 1:length(savedInputs)) { 
      session$sendInputMessage(inputIDs[i],  list(value=inputvalues[[i]]) )
    }
  })
  
  observeEvent(input$save.settings,{ 
    saveRDS(reactiveValuesToList(input), file = 'runs/lastinputs.RDS')

    project_path<-file.path("TutorialEx_MarkdownSetup")
    variant_name<-input$run.name
    input_raster_dir<-"runs/data/01_biodiversity_features/02_habitats/forest"
    
    create_zproject_rev(name="runs",dir=".",variants=variant_name,
                    spp_template_dir=input_raster_dir,spp_file_pattern="*.tif",   
                    overwrite=T,
                    weight=c(input$targ1,input$targ2,input$targ3,input$targ4))

    dat<-list("Settings" = list("removal rule"=2, "warp factor"=50, "edge removal"=1, 
                            "use planning unit layer"=0, "use cost"=0, 
                            "use mask"=0, #"mask file"=act.lyrs[act.lyrs[,'label']==input$action.sel,"masklyr"],
                            "use tree connectivity"=0))
    write_dat(dat,paste0("./runs/",variant_name,"/",variant_name,".dat"),overwrite=T)
    
    output$save.confirm <- renderText({"Your settings have been saved. Now run it!"})
  })  
  
  # Run Zonation from shell upon user input
  observeEvent(input$run.tool,{
    project_path<-file.path("runs")
    variant_name<-input$run.name

    if (file.exists(paste0("./runs/",variant_name,".bat"))) {
      print(getwd())
      shell.exec(paste0("runs\\",   #relative path issue? opening shell then not finding file?
                        variant_name,".bat"))
      output$please.wait<-renderText({paste("Please wait while your prioritization is run. Your results will appear below and will be saved to 'runs/",input$run.name,"'.",sep="")})
    }
    else{
      output$save.error <- renderText({"You must first select and save your run settings"})
    }
  })
```

***




